## No Augmentation

About the technique
- `ts_aug_none()` leaves the training windows unchanged.
- Its value is methodological: it gives you the baseline needed to judge whether synthetic windows are actually helping the predictor.

Didactic goal: establish the reference case before testing any augmentation strategy.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Time series augmentation - none (identity)

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


We load the example series that will be used throughout the demonstration.

```{r}
# Series for study

data(tsd)
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

The next step organizes the series into sliding windows, which is the tabular representation used by the later transformations and models.

```{r}
# Sliding windows

sw_size <- 10
xw <- ts_data(tsd$y, sw_size)
```

Now we augmentation (none).

```{r}
# Augmentation (none)

augment <- ts_aug_none()
set_example_seed()
augment <- fit(augment, xw)
xa <- transform(augment, xw)
idx <- attr(xa, "idx")
ts_head(xa)
```

This plot overlays the original and augmented windows so you can see how the transformation changes the local shape.

```{r}
# Plot a few representative windows on the lag axis
compare_rows <- seq_len(min(nrow(xa), 6))
comparison <- do.call(
  rbind,
  lapply(compare_rows, function(row_id) {
    source_row <- idx[row_id]
    rbind(
      data.frame(lag = seq_len(sw_size), value = as.numeric(xw[source_row, 1:sw_size]), series = "original", sample = paste("window", source_row)),
      data.frame(lag = seq_len(sw_size), value = as.numeric(xa[row_id, 1:sw_size]), series = "augmented", sample = paste("window", source_row))
    )
  })
)

ggplot(comparison, aes(x = lag, y = value, color = series, group = series)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.2) +
  facet_wrap(~ sample, ncol = 3) +
  theme_minimal(base_size = 14)
```

References
- H. I. Fawaz, G. Forestier, J. Weber, L. Idoumghar, and P.-A. Muller (2019). Deep learning for time series classification: A review. Data Mining and Knowledge Discovery, 33, 917–963.

