## Wormhole Augmentation

About the technique
- Wormhole augmentation replaces selected lags with older values from the same series.
- The idea is to create plausible alternative histories that still respect the temporal nature of the data.

Didactic goal: understand a more structural augmentation strategy than simple noise injection.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Time series augmentation - wormhole

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


We load the example series that will be used throughout the demonstration.

```{r}
# Series for study

data(tsd)
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

The next step organizes the series into sliding windows, which is the tabular representation used by the later transformations and models.

```{r}
# Sliding windows

sw_size <- 10
xw <- ts_data(tsd$y, sw_size)
```

Now we augmentation (wormhole).

```{r}
# Augmentation (wormhole)

augment <- ts_aug_wormhole()
set_example_seed()
augment <- fit(augment, xw)
xa <- transform(augment, xw)
idx <- attr(xa, "idx")
ts_head(xa)
```

This plot overlays the original and augmented windows so you can see how the transformation changes the local shape.

```{r}
# Plot a few representative windows on the lag axis
aug_rows <- (nrow(xw) + 1):min(nrow(xa), nrow(xw) + 6)
comparison <- do.call(
  rbind,
  lapply(aug_rows, function(row_id) {
    source_row <- idx[row_id]
    rbind(
      data.frame(lag = seq_len(sw_size), value = as.numeric(xw[source_row, 1:sw_size]), series = "original", sample = paste("window", source_row)),
      data.frame(lag = seq_len(sw_size), value = as.numeric(xa[row_id, 1:sw_size]), series = "augmented", sample = paste("window", source_row))
    )
  })
)

ggplot(comparison, aes(x = lag, y = value, color = series, group = series)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.2) +
  facet_wrap(~ sample, ncol = 3) +
  theme_minimal(base_size = 14)
```

References
- H. I. Fawaz, G. Forestier, J. Weber, L. Idoumghar, and P.-A. Muller (2019). Deep learning for time series classification: A review. Data Mining and Knowledge Discovery, 33, 917–963.

