---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# daltoolbox

<!-- badges: start -->
![GitHub Repo stars](https://img.shields.io/github/stars/cefet-rj-dal/daltoolbox?logo=Github)
<!-- badges: end -->

The goal of DAL Toolbox is to provide a series data analytics functions organized as a framework. 
It supports data preprocessing, classification, regression, clustering, and time series prediction functions. 

## Installation

The latest version of DAL Toolbox at CRAN is available at:
<https://cran.r-project.org/web/packages/daltoolbox/index.html/>

You can install the stable version of DAL Toolbox from CRAN with:

``` r
install.packages("daltoolbox")
```

You can install the development version of DAL Toolbox from [GitHub](https://github.com/) with:

``` r
library(devtools)
devtools::install_github("cefet-rj-dal/daltoolbox", force=TRUE, dependencies=FALSE, upgrade="never")
```

## Examples

The DAL Toolbox examples are made available at:
<https://nbviewer.org/github/cefet-rj-dal/daltoolbox-examples/tree/main/>

A demo video is provided at:
<https://eic.cefet-rj.br/~dal/leveraging-experiment-lines-to-data-analytics/>


The examples are organized according to general (data preprocessing),
clustering, classification, regression, and time series functions.


```{r example}
library(daltoolbox)
## loading DAL Toolbox
```

