The one-dimensional Gabor filter combines a Gaussian envelope with a cosine carrier. In time series smoothing, this creates a local convolution kernel that can attenuate unwanted variation while retaining structure around a selected central frequency.

Notes:
- `window_size` controls the width of the Gaussian envelope in observations.
- `central_frequency` is expressed in cycles per observation and must be between 0 and 0.5. Values close to zero behave more like Gaussian smoothing, while larger values emphasize oscillatory local structure.
- The implementation uses a centered window, so it is intended for offline preprocessing rather than strictly causal forecasting.
- Reflected boundary padding is used so the filtered series keeps the same length as the input.

Objectives: The Gabor filter applies localized frequency-aware smoothing to a time series using a finite symmetric kernel.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Filter - Gabor

# Install tspredit if needed
#install.packages("tspredit")
```

We load the packages required by this example.

```{r}
# Load packages
library(daltoolbox)
library(tspredit)
```

This chunk prepares a noisy seasonal series.

```{r}
set_example_seed()
x <- seq(0, 6 * pi, length.out = 120)
y <- sin(x) + 0.3 * sin(8 * x) + rnorm(length(x), 0, 0.12)
```

We plot the data here so the effect of the transformation can be compared visually.

```{r}
library(ggplot2)
# Visualize noisy input
plot_ts(y = y) + theme(text = element_text(size = 16))
```

We now apply the Gabor filter.

```{r}
# Apply Gabor smoothing
filter <- ts_fil_gabor(window_size = 3, central_frequency = 0.05)
filter <- daltoolbox::fit(filter, y)
yhat <- transform(filter, y)

# Compare original vs filtered series
plot_ts_pred(y = y, yadj = yhat) + theme(text = element_text(size = 16))
```

References
- D. Gabor (1946). Theory of communication. Journal of the Institution of Electrical Engineers - Part III: Radio and Communication Engineering, 93, 429-457.
