Splines filter: Smoothing splines minimize a penalized least squares criterion that trades fidelity to the data against curvature of the function. The smoothing parameter controls the penalty on the integrated squared second derivative, yielding a flexible nonparametric smoother.

Objective: Apply cubic smoothing splines to reduce noise by controlling roughness via the spar parameter.

```{r}
# Filter - Spline

# Installing the package (if needed)
#install.packages("tspredit")
```

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


```{r}
# Series for study with artificial noise and spikes

data(tsd)
y <- tsd$y
noise <- rnorm(length(y), 0, sd(y)/10)
spike <- rnorm(1, 0, sd(y))
tsd$y <- tsd$y + noise
tsd$y[10] <- tsd$y[10] + spike
tsd$y[20] <- tsd$y[20] + spike
tsd$y[30] <- tsd$y[30] + spike
```

```{r}
library(ggplot2)
# Noisy series visualization
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

```{r}
# Applying the Spline filter

filter <-  ts_fil_spline(spar = 0.5)
filter <- fit(filter, tsd$y)
y <- transform(filter, tsd$y)
plot_ts_pred(y=tsd$y, yadj=y) + theme(text = element_text(size=16))
```

References
- G. Wahba (1990). Spline Models for Observational Data. SIAM.
