## Custom Normalization

The goal of this example is to show how to create a custom normalization strategy that remains compatible with the `tspredit` forecasting pipeline.

The built-in adaptive normalization in the package centers each window by subtracting a moving average and then applies global rescaling. Here we implement a didactic variant inspired by the original adaptive-normalization idea: divide by the moving average first, then apply global min-max scaling. This is useful when we want the model to focus on relative level instead of absolute magnitude.

Because divisive normalization is easier to interpret on strictly positive series, this example uses the GDP benchmark data.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# installation
# install.packages(c("tspredit", "daltoolbox"))

library(daltoolbox)
library(tspredit)
```

This chunk defines the custom class and the S3 methods needed to integrate it with the `tspredit` workflow.

```{r}
ts_norm_an_div_custom <- function(outliers = daltoolbox::outliers_boxplot(),
                                  nw = 0,
                                  eps = 1e-8) {
  obj <- daltoolbox::dal_transform()
  obj$outliers <- outliers
  obj$nw <- nw
  obj$eps <- eps
  obj$an_mean <- mean
  obj$ma <- function(obj, data, func) {
    data <- unclass(as.matrix(data))
    if (obj$nw != 0) {
      cols <- ncol(data) - ((obj$nw - 1):0)
      data <- data[, cols, drop = FALSE]
    }
    apply(data, 1, func, na.rm = TRUE)
  }
  class(obj) <- append("ts_norm_an_div_custom", class(obj))
  obj
}

fit.ts_norm_an_div_custom <- function(obj, data, ...) {
  input <- data[, 1:(ncol(data) - 1)]
  an <- obj$ma(obj, input, obj$an_mean)
  denom <- ifelse(abs(an) < obj$eps, obj$eps, an)
  data <- sweep(data, 1, denom, "/")

  if (!is.null(obj$outliers)) {
set_example_seed()
    out <- fit(obj$outliers, data)
    data <- transform(out, data)
  }

  obj$gmin <- min(data)
  obj$gmax <- max(data)
  obj
}

transform.ts_norm_an_div_custom <- function(obj, data, x = NULL, ...) {
  if (!is.null(x)) {
    denom <- attr(data, "an_denom")
    x <- x / denom
    x <- (x - obj$gmin) / (obj$gmax - obj$gmin)
    return(x)
  }

  an <- obj$ma(obj, data, obj$an_mean)
  denom <- ifelse(abs(an) < obj$eps, obj$eps, an)
  data <- sweep(data, 1, denom, "/")
  data <- (data - obj$gmin) / (obj$gmax - obj$gmin)
  attr(data, "an_mean") <- an
  attr(data, "an_denom") <- denom
  data
}

inverse_transform.ts_norm_an_div_custom <- function(obj, data, x = NULL, ...) {
  denom <- attr(data, "an_denom")

  if (!is.null(x)) {
    x <- x * (obj$gmax - obj$gmin) + obj$gmin
    x <- x * denom
    return(x)
  }

  data <- data * (obj$gmax - obj$gmin) + obj$gmin
  data <- data * denom
  attr(data, "an_denom") <- denom
  data
}

registerS3method("fit", "ts_norm_an_div_custom", fit.ts_norm_an_div_custom)
registerS3method("transform", "ts_norm_an_div_custom", transform.ts_norm_an_div_custom)
registerS3method("inverse_transform", "ts_norm_an_div_custom", inverse_transform.ts_norm_an_div_custom)
```

We first inspect the transformed windows to see how the custom normalizer changes scale.

```{r}
data(gdp)
series <- gdp$usa_gdp

ts <- ts_data(series, 10)
normalizer <- ts_norm_an_div_custom(nw = 3)
set_example_seed()
normalizer <- fit(normalizer, ts)
tst <- transform(normalizer, ts)

ts_head(tst, 3)
```

The next step shows that the custom normalizer can be plugged directly into a predictive model.

```{r}
samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test <- ts_projection(samp$test)
```

This chunk configures the model and fits it on the training data prepared earlier.

```{r}
model <- ts_mlp(
  preprocess = ts_norm_an_div_custom(nw = 3),
  input_size = 4,
  size = 4,
  decay = 0,
  maxit = 1000
)

set_example_seed()
model <- fit(model, x = io_train$input, y = io_train$output)
prediction <- as.vector(predict(model, x = io_test$input[1:1, ], steps_ahead = 5))
evaluate(model, as.vector(io_test$output), prediction)$metrics
```

This example shows that a custom normalizer in `tspredit` needs to provide the same practical contract as the built-in preprocessors: a fitted state, a forward transformation, and an inverse transformation for predictions.

References
- Ogasawara, E., Martinez, L. C., De Oliveira, D., Zimbrão, G., Pappa, G. L., Mattoso, M. (2010). Adaptive Normalization: A novel data normalization approach for non-stationary time series.
