## Build Multivariate Time-Series Data

About the concept
- `ts_data_mv()` is the multivariate entry point of `tspredit` 2.0.
- It does not build lagged windows immediately. Its job is to preserve the temporal alignment between the target `y` and the auxiliary variables `x1, ..., xn`.

Didactic goal: understand how the multivariate workflow starts before any forecasting model is trained.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Build multivariate time-series data

# Installing the package (if needed)
# install.packages("tspredit")
```

We begin by loading the packages used in the example.

```{r}
library(daltoolbox)
library(tspredit)
```

We now use package benchmarks already distributed with `tspredit`. We extract:

- the daily maximum load from `EUNITE.Loads` as the target `y`
- the weekday code from `EUNITE.Reg` as `x1`
- a weekend indicator derived from `EUNITE.Reg` as `x2`

```{r}
data(EUNITE.Loads)
data(EUNITE.Reg)

if (!is.null(attr(EUNITE.Loads, "url"))) {
  EUNITE.Loads <- loadfulldata(EUNITE.Loads)
}
if (!is.null(attr(EUNITE.Reg, "url"))) {
  EUNITE.Reg <- loadfulldata(EUNITE.Reg)
}

load_cols <- setdiff(names(EUNITE.Loads), "split")
y <- apply(EUNITE.Loads[, load_cols, drop = FALSE], 1, max)
x1 <- as.numeric(EUNITE.Reg$Weekday)
x2 <- as.numeric(EUNITE.Reg$Weekday %in% c(1, 7))
```

We now build the aligned multivariate object. Every row corresponds to the same
time instant for `y`, `x1`, and `x2`.

```{r}
mv <- ts_data_mv(
  data.frame(
    y = y,
    x1 = x1,
    x2 = x2
  ),
  y = "y"
)

ts_head(mv, 5)
```

The object keeps the target and the auxiliary variable names as metadata.

```{r}
attr(mv, "y")
attr(mv, "x")
attr(mv, "variables")
```

The next step in most workflows is a time-aware split.

```{r}
samp <- ts_sample(mv, test_size = 5)
ts_head(samp$train, 3)
ts_head(samp$test, 3)
```

What this example shows
- `ts_data_mv()` is the aligned multivariate representation used by the rest of the multivariate workflow.
- The object preserves the semantic distinction between the target and the auxiliary variables.
- `ts_sample()` works on `ts_data_mv()` just as it does on the univariate workflow, but without losing temporal alignment.
