## Train and Test Splits for Singular Multivariate Data

About the concept
- The singular multivariate branch works with aligned observations, that is, with `ts_data_mv(..., sw = 1)`.
- This is the natural input for the new `ts_reg_mv` family, which includes `ts_lm_mv()`, `ts_arimax()`, and `ts_var()`.

Didactic goal: inspect how the aligned multivariate object is split in time before fitting singular multivariate models.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Train and test splits for singular multivariate data

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

For the singular branch, it is useful to start from a compact synthetic
multivariate system built from the benchmark series `tsd$y`.

```{r}
data(tsd)

x1 <- c(tsd$y[-1], tail(tsd$y, 1))
x2 <- stats::filter(tsd$y, rep(1/3, 3), sides = 1)
x2[is.na(x2)] <- tsd$y[is.na(x2)]

mv <- ts_data_mv(
  data.frame(y = tsd$y, x1 = x1, x2 = as.numeric(x2)),
  y = "y"
)

ts_head(mv, 5)
```

Because this branch operates directly on synchronized observations, the split is
performed on the aligned object itself.

```{r}
samp <- ts_sample(mv, test_size = 5)

ts_head(samp$train, 3)
ts_head(samp$test, 3)
```

The target remains explicit, but the future auxiliary values are also available
in the held-out partition. This matters because singular multivariate models may
use them in two different ways:

- directly, when the future auxiliaries are already known
- indirectly, when the auxiliaries must be forecast by their own univariate models

```{r}
tail(samp$train$y, 3)
as.data.frame(samp$test)[, c("x1", "x2")]
```

What this example shows
- `ts_sample()` should be applied directly to the aligned `ts_data_mv()` object in the singular multivariate branch.
- The test partition keeps the future rows of `x1, ..., xn`, which can be supplied directly to `ts_lm_mv()` or `ts_arimax()`.
- This branch is conceptually different from the sliding-window branch, where the split happens after the lagged materialization.
