## Target-Centered Multivariate Linear Regression

About the method
- `ts_lm_mv()` is the simplest model in the singular multivariate branch.
- It keeps one target `y` as the main forecasting objective and expresses its relation with the auxiliary variables through a regression formula.

Didactic goal: show how the formula-based style from `daltoolbox` enters the multivariate time-series workflow of `tspredit`.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Target-centered multivariate linear regression

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

We reuse a compact synthetic multivariate system built from the benchmark series
`tsd$y`.

```{r}
data(tsd)

x1 <- c(tsd$y[-1], tail(tsd$y, 1))
x2 <- stats::filter(tsd$y, rep(1/3, 3), sides = 1)
x2[is.na(x2)] <- tsd$y[is.na(x2)]

mv <- ts_data_mv(
  data.frame(y = tsd$y, x1 = x1, x2 = as.numeric(x2)),
  y = "y"
)

samp <- ts_sample(mv, test_size = 5)
```

The formula makes the relation between the target and the auxiliaries explicit.

```{r}
model <- ts_lm_mv(formula = y ~ x1 + x2)
model <- fit(model, samp$train)
```

In this first singular example, we assume that the future auxiliaries are known
and therefore use the held-out rows directly during prediction.

```{r}
pred_1 <- predict(model, x = samp$test, steps_ahead = 1)
pred_1
```

```{r}
pred_5 <- predict(model, x = samp$test, steps_ahead = 5)
pred_5
```

When we ask for the complete object, the target forecast is returned together
with the auxiliary values that were used in the horizon.

```{r}
pred_all <- predict(model, x = samp$test, steps_ahead = 5, return_all = TRUE)
pred_all
```

The target values remain available for direct evaluation.

```{r}
ev_test <- evaluate(model, tail(samp$test$y, 5), pred_5)
ev_test$metrics
```

What this example shows
- `ts_lm_mv()` is the most direct entry point into the singular multivariate branch.
- The `formula` argument makes the structural relation between `y` and the auxiliary variables explicit.
- When future auxiliary values are known, singular multivariate prediction can use the held-out aligned rows directly.
