## Train and Test Splits for Multivariate Data

About the concept
- In the multivariate workflow, train/test splitting happens on the aligned object before forecasting.
- The split must preserve the temporal synchronization between `y` and all auxiliary variables.

Didactic goal: inspect how the aligned multivariate data and the lagged materialization behave before and after the time-aware split.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Train and test splits for multivariate data

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

```{r}
data(EUNITE.Loads)
data(EUNITE.Reg)

if (!is.null(attr(EUNITE.Loads, "url"))) {
  EUNITE.Loads <- loadfulldata(EUNITE.Loads)
}
if (!is.null(attr(EUNITE.Reg, "url"))) {
  EUNITE.Reg <- loadfulldata(EUNITE.Reg)
}

load_cols <- setdiff(names(EUNITE.Loads), "split")
y <- apply(EUNITE.Loads[, load_cols, drop = FALSE], 1, max)
x1 <- as.numeric(EUNITE.Reg$Weekday)
x2 <- as.numeric(EUNITE.Reg$Weekday %in% c(1, 7))

mv <- ts_data_mv(
  data.frame(y = y, x1 = x1, x2 = x2),
  y = "y"
)
```

We first split the aligned multivariate object.

```{r}
samp <- ts_sample(mv, test_size = 5)

ts_head(samp$train, 3)
ts_head(samp$test, 3)
```

The same split can then be materialized into lagged windows separately for training and test.

```{r}
train_windows <- ts_window_mv(samp$train, window_size = 7)
test_windows <- ts_window_mv(samp$test, window_size = 5)

ts_head(train_windows, 3)
ts_head(test_windows, 1)
```

This makes the temporal boundary explicit: the learner sees lagged windows built from the training portion only, while the held-out tail remains untouched for forecasting and evaluation.

What this example shows
- `ts_sample()` keeps the multivariate alignment intact.
- Window materialization can be applied independently to train and test partitions.
- The forecasting workflow remains time-aware from the raw data to the lagged representation.
