## Target-Centered ARIMAX with Auxiliary Forecasting

About the method
- `ts_arimax()` extends the raw-series ARIMA branch to aligned multivariate observations.
- It keeps one target `y`, uses the auxiliary series as external regressors, and can rely on univariate support models when future auxiliaries are unknown.

Didactic goal: show how the singular multivariate branch remains target-centered while still forecasting the auxiliaries needed along the horizon.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Target-centered ARIMAX with auxiliary forecasting

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

We again start from a small aligned multivariate system derived from `tsd$y`.

```{r}
data(tsd)

x1 <- c(tsd$y[-1], tail(tsd$y, 1))
x2 <- stats::filter(tsd$y, rep(1/3, 3), sides = 1)
x2[is.na(x2)] <- tsd$y[is.na(x2)]

mv <- ts_data_mv(
  data.frame(y = tsd$y, x1 = x1, x2 = as.numeric(x2)),
  y = "y"
)

samp <- ts_sample(mv, test_size = 5)
```

This notebook follows the same singular-experiment structure as the linear
regression example, but now the target model is `ARIMAX` and the auxiliary
paths are forecast by support models. Before that, we keep a direct univariate
`ARIMA` baseline on the target.

```{r}
y_ts <- ts_data(tsd$y, 1)
y_samp <- ts_sample(y_ts, test_size = 5)

baseline <- ts_arima()
baseline <- fit(baseline, y_samp$train)
pred_base <- predict(baseline, y_samp$test[1, ], steps_ahead = 5)
ev_base <- evaluate(baseline, as.vector(y_samp$test), as.vector(pred_base))
ev_base$metrics
```

We now build the multivariate singular model. The target is learned with
`ARIMAX`, while the future auxiliary paths are generated by univariate
`ARIMA` submodels.

```{r}
model <- ts_arimax(
  models_x = list(
    x1 = ts_arima(),
    x2 = ts_arima()
  )
)

model <- fit(model, samp$train)
```

```{r}
pred_1 <- predict(model, steps_ahead = 1)
pred_1
```

```{r}
pred_5 <- predict(model, steps_ahead = 5)
pred_5
```

```{r}
attr(pred_5, "system")
```

```{r}
ev_test <- evaluate(model, tail(samp$test$y, 5), pred_5)
rbind(
  ARIMA_y = ev_base$metrics,
  ARIMAX_mv = ev_test$metrics
)
```

What this example shows
- `ts_arimax()` is the singular multivariate extension of the raw-series `ARIMA` branch.
- The target remains the semantic center of the workflow, but the forecast horizon also requires trajectories for the auxiliary variables.
- These future auxiliary paths can be generated by their own univariate support models.
