## Deterministic Auxiliary Models

About the concept
- In a target-centered multivariate workflow, not every auxiliary variable needs a generic learner.
- Some variables follow an explicit law of formation, so the correct object is a deterministic predictor rather than a statistical model.

Didactic goal: show how the `ts_deterministic()` family fits naturally inside the multivariate architecture as the pipeline for structured auxiliary variables.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Deterministic auxiliary models in multivariate forecasting

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

We use the same multivariate benchmark structure adopted throughout this
section:

- `y`: daily maximum load
- `x1`: weekday code
- `x2`: weekend indicator

```{r}
data(EUNITE.Loads)
data(EUNITE.Reg)

if (!is.null(attr(EUNITE.Loads, "url"))) {
  EUNITE.Loads <- loadfulldata(EUNITE.Loads)
}
if (!is.null(attr(EUNITE.Reg, "url"))) {
  EUNITE.Reg <- loadfulldata(EUNITE.Reg)
}

load_cols <- setdiff(names(EUNITE.Loads), "split")
y <- apply(EUNITE.Loads[, load_cols, drop = FALSE], 1, max)
x1 <- as.numeric(EUNITE.Reg$Weekday)
x2 <- as.numeric(EUNITE.Reg$Weekday %in% c(1, 7))

mv <- ts_data_mv(
  data.frame(y = y, x1 = x1, x2 = x2),
  y = "y"
)

samp <- ts_sample(mv, test_size = 5)
```

The point here is not yet the full forecasting experiment. The point is to
decide what kind of univariate object makes sense for each auxiliary variable.

`x1` is a periodic calendar code. It is more coherent to model it with a
deterministic periodic rule than with a neural network or a random forest.

```{r}
model_x1 <- ts_mv_spec(
  ts_deterministic("periodic", period = 7)
)

class(model_x1$model)
```

`x2` is a binary weekend flag. It also follows a periodic law of formation, so
it can use the same deterministic family.

```{r}
model_x2 <- ts_mv_spec(
  ts_periodic(7)
)

class(model_x2$model)
```

For slowly changing or operationally fixed auxiliary variables, persistence is
often enough.

```{r}
model_persist <- ts_mv_spec(
  ts_persist()
)

class(model_persist$model)
```

These objects are regular multivariate specifications. They plug into the same
contract as any other auxiliary learner.

```{r}
model <- ts_regsw_mv(
  model_y = ts_mv_spec(
    ts_mlp(ts_norm_an(), input_size = 4, size = 4, decay = 0),
    variables = c("y", "x1", "x2")
  ),
  models_x = list(
    x1 = model_x1,
    x2 = model_x2
  ),
  window_size = 7
)

class(model)
```

We can already fit the composed object and inspect the auxiliary recursive path.

```{r}
set_example_seed()
model <- fit(model, samp$train)
pred_all <- predict(model, steps_ahead = 5, return_all = TRUE)

pred_all$x
```

What this example shows
- `ts_deterministic()` is not a separate multivariate subsystem. It is a family of univariate objects that plugs into the multivariate orchestration.
- `ts_periodic()` and `ts_persist()` remain useful wrappers when the intent should be explicit.
- Deterministic auxiliary variables should usually be modeled by their law of formation, not by generic learners.
- This choice is best taught inside the multivariate workflow, because that is where these auxiliary variables become operationally important.
