## Target-Centered Vector Autoregression

About the method
- `ts_var()` models the multivariate system jointly.
- Even so, `tspredit` keeps one target `y` as the main variable of interest and returns its forecast by default.

Didactic goal: show how a system model can coexist with a target-centered interface.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Target-centered vector autoregression

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

We use the same aligned multivariate system as in the previous singular
examples, so the only change is the model family.

```{r}
data(tsd)

x1 <- c(tsd$y[-1], tail(tsd$y, 1))
x2 <- stats::filter(tsd$y, rep(1/3, 3), sides = 1)
x2[is.na(x2)] <- tsd$y[is.na(x2)]

mv <- ts_data_mv(
  data.frame(y = tsd$y, x1 = x1, x2 = as.numeric(x2)),
  y = "y"
)

samp <- ts_sample(mv, test_size = 5)
```

The vector autoregression is fitted on the whole aligned system. Here, the
target remains `y`, but the system forecast includes `x1` and `x2` as well.

```{r}
model <- ts_var(p_max = 3)
model <- fit(model, samp$train)
```

```{r}
pred_5 <- predict(model, steps_ahead = 5)
pred_5
```

```{r}
pred_all <- predict(model, steps_ahead = 5, return_all = TRUE)
pred_all
```

The target can still be evaluated exactly as in the target-centered models.

```{r}
ev_test <- evaluate(model, tail(samp$test$y, 5), pred_5)
ev_test$metrics
```

What this example shows
- `ts_var()` models the multivariate system jointly, not just one regression equation for `y`.
- Even in that systemic setting, the `tspredit` interface can keep a distinguished target variable.
- `return_all = TRUE` is the key to inspect the full system forecast while preserving the target-centered default behavior.
