## Train and Test Splits for Windowed Multivariate Data

About the concept
- In sliding-window forecasting workflows, train/test splitting should happen after the lagged representation is materialized.
- The split must preserve the temporal synchronization between `y` and all auxiliary variables.

Didactic goal: inspect how the aligned multivariate data becomes lagged windows and why the split should happen at that stage.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Train and test splits for windowed multivariate data

# Installing the package (if needed)
# install.packages("tspredit")
```

```{r}
library(daltoolbox)
library(tspredit)
```

```{r}
data(EUNITE.Loads)
data(EUNITE.Reg)

if (!is.null(attr(EUNITE.Loads, "url"))) {
  EUNITE.Loads <- loadfulldata(EUNITE.Loads)
}
if (!is.null(attr(EUNITE.Reg, "url"))) {
  EUNITE.Reg <- loadfulldata(EUNITE.Reg)
}

load_cols <- setdiff(names(EUNITE.Loads), "split")
y <- apply(EUNITE.Loads[, load_cols, drop = FALSE], 1, max)
x1 <- as.numeric(EUNITE.Reg$Weekday)
x2 <- as.numeric(EUNITE.Reg$Weekday %in% c(1, 7))

mv <- ts_data_mv(
  data.frame(y = y, x1 = x1, x2 = x2),
  y = "y"
)
```

We first materialize the aligned multivariate object into lagged windows.

```{r}
windows <- ts_data_mv(mv, sw = 7)

ts_head(windows, 3)
```

We then split the lagged representation into train and test partitions.

```{r}
samp <- ts_sample(windows, test_size = 10)

ts_head(samp$train, 3)
ts_head(samp$test, 3)
```

This is the important detail inherited from the univariate workflow: the first
rows of the test partition already carry lagged values whose context comes from
the end of the training period. If we split the raw aligned series first and
only then materialize the windows, we lose that boundary information.

What this example shows
- `ts_data_mv(..., sw > 1)` is the multivariate bridge from aligned observations to lagged forecasting inputs.
- `ts_sample()` should be applied after the lagged representation is materialized when the goal is a sliding-window train/test split.
- The forecasting workflow remains time-aware from the raw multivariate data to the final train/test window partitions.
