## Objective

This notebook introduces `tsd`, the smallest dataset distributed with `tspredit`. The goal is to inspect its structure and plot the synthetic target series used throughout the introductory examples.

## Method at a glance

This is a dataset-orientation notebook. It does not fit a model. Instead, it inspects the packaged object so the reader understands why `tsd` is the didactic entry point of the package.

## What you will do

- load `tsd`
- inspect its dimensions and columns
- summarize the variables
- plot the target series

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
library(tspredit)
library(ggplot2)
```

```{r}
data(tsd)
cat("Dataset: tsd\n")
cat("Rows:", nrow(tsd), "\n")
cat("Columns:", paste(names(tsd), collapse = ", "), "\n")
head(tsd)
```

```{r}
summary(tsd)
```

```{r}
plot_ts(x = tsd$x, y = tsd$y) + theme(text = element_text(size = 16))
```

## References

- Generated for package examples and documentation.
