## Objective

This notebook introduces `EUNITE.Reg`, the regressor table packaged with the EUNITE competition data.

## Method at a glance

The goal is to inspect the categorical and calendar-style variables that complement the load and temperature series.

## What you will do

- load `EUNITE.Reg`
- inspect dimensions and column names
- preview the first rows
- summarize the regressors

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
library(tspredit)
```

```{r}
expand_dataset <- function(x) {
  url <- attr(x, "url")
  if (is.null(url) || !nzchar(url)) x else loadfulldata(x)
}
```

```{r}
data(EUNITE.Reg)
EUNITE.Reg <- expand_dataset(EUNITE.Reg)
cat("Dataset: EUNITE.Reg\n")
cat("Rows:", nrow(EUNITE.Reg), "\n")
cat("Columns:", paste(names(EUNITE.Reg), collapse = ", "), "\n")
head(EUNITE.Reg)
```

```{r}
summary(EUNITE.Reg)
```

## References

- Chen, B.-J., Chang, M.-W., and Lin, C.-J. (2004). Load forecasting using support vector machines: a study on EUNITE competition 2001.
