## Objective

This notebook introduces `NN3`, the monthly forecasting competition archive.

## Method at a glance

The notebook inspects the wide data-frame representation, where each column corresponds to one series in the archive.

## What you will do

- load `NN3`
- inspect dimensions and column names
- preview the first rows
- plot one representative series

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
library(tspredit)
```

```{r}
expand_dataset <- function(x) {
  url <- attr(x, "url")
  if (is.null(url) || !nzchar(url)) x else loadfulldata(x)
}
```

```{r}
data(NN3)
NN3 <- expand_dataset(NN3)
cat("Dataset: NN3\n")
cat("Rows:", nrow(NN3), "\n")
cat("Columns:", ncol(NN3), "\n")
head(names(NN3))
head(NN3[, 1:4])
```

```{r}
ts.plot(NN3[[1]], ylab = "Value", xlab = "Month", main = names(NN3)[1])
```

## References

- Crone, S. F., Hibon, M., and Nikolopoulos, K. (2011). Advances in forecasting with neural networks? Empirical evidence from the NN3 competition on time series prediction.
