## Objective

This notebook introduces `stocks`, the packaged B3 stock-market collection.

## Method at a glance

The notebook inspects the list-of-data-frames structure and previews one ticker entry with its OHLCV fields.

## What you will do

- load `stocks`
- inspect the available tickers
- preview the structure of the first ticker
- plot the closing-price series

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
library(tspredit)
```

```{r}
expand_dataset <- function(x) {
  url <- attr(x, "url")
  if (is.null(url) || !nzchar(url)) x else loadfulldata(x)
}
```

```{r}
data(stocks)
stocks <- expand_dataset(stocks)
cat("Dataset: stocks\n")
cat("Tickers available:", length(stocks), "\n")
head(names(stocks))
first_ticker <- names(stocks)[1]
first_series <- stocks[[first_ticker]]
head(first_series)
```

```{r}
ts.plot(first_series$close, ylab = "Close", xlab = "Index", main = paste(first_ticker, "close"))
```

## References

- B3 historical trading data.
