## No Normalization

About the technique

- `ts_norm_none()` leaves the numerical scale unchanged.
- Its role is to establish whether a predictor already works well on the raw scale before introducing transformations.

Didactic goal: use the raw series as the baseline for judging later normalization strategies.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# No normalization

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
library(daltoolbox)
library(tspredit)
library(ggplot2)
```

We load the example series that will be used throughout the demonstration.

```{r}
data(tsd)
```

The first plot shows the original series. This is the common visual reference
for all normalization examples in this folder.

```{r}
plot_ts(x = tsd$x, y = tsd$y) + theme(text = element_text(size = 16))
```

The next step organizes the series into sliding windows, which is the tabular
representation used by the later transformations and models.

```{r}
sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[, 10])
```

We now keep the values unchanged and compare the supervised target column
(`t0`) before and after the transformation.

```{r}
preproc <- ts_norm_none()
set_example_seed()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[, 10])

compare_t0 <- rbind(
  data.frame(idx = seq_len(nrow(ts)), value = as.vector(ts[, ncol(ts)]), series = "original t0"),
  data.frame(idx = seq_len(nrow(tst)), value = as.vector(tst[, ncol(tst)]), series = "transformed t0")
)

ggplot(compare_t0, aes(x = idx, y = value, color = series)) +
  geom_line(linewidth = 0.7) +
  theme_minimal(base_size = 14)
```

What to observe

- The two curves should overlap completely.
- This file is the baseline reference for the other normalization examples.

References

- C. M. Bishop (2006). Pattern Recognition and Machine Learning. Springer.
