## Global Min-Max Normalization

About the technique

- Global min-max normalization rescales the training values to a fixed interval, usually `[0, 1]`.
- It preserves relative ordering while making magnitudes easier for many learning algorithms to handle.

Didactic goal: inspect the simplest global scaling strategy used in forecasting pipelines.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Global Min-Max Normalization

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
library(daltoolbox)
library(tspredit)
library(ggplot2)
```

We load the example series that will be used throughout the demonstration.

```{r}
data(tsd)
```

The first plot shows the original series. This is the common visual reference
for all normalization examples in this folder.

```{r}
plot_ts(x = tsd$x, y = tsd$y) + theme(text = element_text(size = 16))
```

The next step organizes the series into sliding windows, which is the tabular
representation used by the later transformations and models.

```{r}
sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[, 10])
```

We now apply global min-max normalization and compare the supervised target
column (`t0`) before and after the transformation.

```{r}
preproc <- ts_norm_gminmax()
set_example_seed()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[, 10])

compare_t0 <- rbind(
  data.frame(idx = seq_len(nrow(ts)), value = as.vector(ts[, ncol(ts)]), series = "original t0"),
  data.frame(idx = seq_len(nrow(tst)), value = as.vector(tst[, ncol(tst)]), series = "transformed t0")
)

plot_ts_pred(
  x = compare_t0[compare_t0$series == "original t0", "idx"],
  y = compare_t0[compare_t0$series == "original t0", "value"],
  yadj = compare_t0[compare_t0$series == "transformed t0", "value"]
) + theme(text = element_text(size = 16))
```

What to observe

- The transformed curve is restricted to a common global interval.
- The ordering of peaks and valleys is preserved, but the original magnitude is removed.

References

- C. M. Bishop (2006). Pattern Recognition and Machine Learning. Springer.
