## Adaptive Subtractive Normalization

About the technique

- Subtractive adaptive normalization removes the adaptive local level from each window.
- It is useful when the relevant signal is a deviation around a moving baseline, especially near zero where divisive normalization can become unstable.
- Within the adaptive-normalization family implemented by `ts_norm_an()`, this corresponds to `operation = "subtract"`.

Didactic goal: see adaptive normalization as a local detrending operator that preserves additive contrasts.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Adaptive Subtractive Normalization

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
library(daltoolbox)
library(tspredit)
library(ggplot2)
```

We load the example series that will be used throughout the demonstration.

```{r}
data(tsd)
```

The first plot shows the original series. This is the common visual reference
for all normalization examples in this folder.

```{r}
plot_ts(x = tsd$x, y = tsd$y) + theme(text = element_text(size = 16))
```

The next step organizes the series into sliding windows, which is the tabular
representation used by the later transformations and models.

```{r}
sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[, 10])
```

We now apply the subtractive version of adaptive normalization and compare the
supervised target column (`t0`) before and after the transformation.

```{r}
preproc <- ts_norm_an(operation = "subtract")
set_example_seed()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[, 10])

compare_t0 <- rbind(
  data.frame(idx = seq_len(nrow(ts)), value = as.vector(ts[, ncol(ts)]), series = "original t0"),
  data.frame(idx = seq_len(nrow(tst)), value = as.vector(tst[, ncol(tst)]), series = "transformed t0")
)

ggplot(compare_t0, aes(x = idx, y = value, color = series)) +
  geom_line(linewidth = 0.7) +
  theme_minimal(base_size = 14)
```

What to observe

- The transformed target behaves like a moving-baseline deviation.
- This is the safest adaptive choice when the local reference level can be close to zero.

References

- Ogasawara, E., Martinez, L. C., De Oliveira, D., Zimbrão, G., Pappa, G. L., Mattoso, M. (2010).
Adaptive Normalization: A novel data normalization approach for non-stationary time series.
Proceedings of the International Joint Conference on Neural Networks (IJCNN).
doi:10.1109/IJCNN.2010.5596746
