## Adaptive Asinh Normalization

About the technique

- Adaptive asinh normalization applies an inverse-hyperbolic-sine contrast around the adaptive local level.
- It is useful when the analyst wants a transformation that is approximately linear near zero and progressively log-like for larger magnitudes.
- Within the adaptive-normalization family implemented by `ts_norm_an()`, this corresponds to `operation = "asinh"`.
- The adaptive reference is estimated on the full supervised window, so the nonlinear contrast is computed over the same complete window that reaches the downstream model.

Didactic goal: understand a smooth nonlinear alternative that connects additive and multiplicative interpretations without a hard switch.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Adaptive Asinh Normalization

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
library(daltoolbox)
library(tspredit)
library(ggplot2)
```

We load the example series that will be used throughout the demonstration.

```{r}
data(tsd)
```

The first plot shows the original series. This is the common visual reference
for all normalization examples in this folder.

```{r}
plot_ts(x = tsd$x, y = tsd$y) + theme(text = element_text(size = 16))
```

The next step organizes the series into sliding windows, which is the tabular
representation used by the later transformations and models.

```{r}
sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[, 10])
```

We now apply the adaptive asinh operator and compare the supervised target
column (`t0`) before and after the transformation.

```{r}
preproc <- ts_norm_an(operation = "asinh", scale = "sd", lambda = 1)
set_example_seed()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[, 10])

compare_t0 <- rbind(
  data.frame(idx = seq_len(nrow(ts)), value = as.vector(ts[, ncol(ts)]), series = "original t0"),
  data.frame(idx = seq_len(nrow(tst)), value = as.vector(tst[, ncol(tst)]), series = "transformed t0")
)

plot_ts_pred(
  x = compare_t0[compare_t0$series == "original t0", "idx"],
  y = compare_t0[compare_t0$series == "original t0", "value"],
  yadj = compare_t0[compare_t0$series == "transformed t0", "value"]
) + theme(text = element_text(size = 16))
```

What to observe

- Near zero, the transformed target stays close to an additive contrast.
- For larger deviations, the transformation becomes progressively more log-like without switching abruptly.

References

- Burbidge, J. B., Magee, L., Robb, A. L. (1988). Alternative Transformations to Handle Extreme Values of the Dependent Variable.
Journal of the American Statistical Association, 83(401), 123-127.
- Bellemare, M. F., Wichman, C. J. (2020). Elasticities and the Inverse Hyperbolic Sine Transformation.
Oxford Bulletin of Economics and Statistics, 82(1), 50-61. doi:10.1111/obes.12325
