## Adaptive Normalization Visual Comparison

This notebook compares the main adaptive normalization operators on synthetic
series designed to stress different theoretical axes of the method.

What you will learn

- How divisive normalization reacts when the adaptive level is close to zero
- How subtractive normalization behaves under changing amplitude
- How `softdivide` interpolates between additive and relative behavior
- How `asinh` changes the geometry of local deviations

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Adaptive Normalization Visual Comparison

# Installing the package (if needed)
#install.packages("tspredit")
```

We load the packages required by this example.

```{r}
library(daltoolbox)
library(tspredit)
library(ggplot2)
```

We build four synthetic series, each emphasizing one regime where the
normalizers behave differently.

```{r}
set_example_seed()

n <- 200
t <- seq_len(n)

synthetic <- list(
  near_zero = 0.03 * sin(t / 5) + rnorm(n, sd = 0.01),
  level_shift = c(rep(0.3, 70), rep(1.2, 60), rep(0.6, 70)) + 0.08 * sin(t / 7),
  heteroscedastic = sin(t / 6) * seq(0.2, 2.2, length.out = n),
  mixed = 0.2 * sin(t / 8) + seq(-0.4, 1.8, length.out = n) + rnorm(n, sd = seq(0.02, 0.25, length.out = n))
)
```

## Comparison Strategy

To keep the comparison fair, every operator uses the same sliding-window view.
We inspect the operators in two complementary ways:

- a fixed anchor window, so we can compare the transformed lag profile directly
- the transformed target dimension (`t0`) across time, which is the signal most
  directly consumed by the supervised pipeline

In the current `ts_norm_an()` contract, the adaptive reference is estimated on
the full supervised window, not only on the lagged inputs. So every curve below
reflects the complete window-wise transformation that the operator applies in
practice.

This separation matters. If we only plot the transformed target across time,
the comparison becomes hard to read because each operator is reacting to a
different local window at each time step. The anchor-window plot removes this
confounding factor and answers a simpler question: how do the operators
transform the same local pattern?

```{r}
ops <- list(
  divide = ts_norm_an(outliers = NULL, operation = "divide"),
  subtract = ts_norm_an(outliers = NULL, operation = "subtract"),
  softdivide = ts_norm_an(outliers = NULL, operation = "softdivide", scale = "sd", lambda = 1),
  asinh = ts_norm_an(outliers = NULL, operation = "asinh", scale = "sd", lambda = 1)
)

anchor_row <- function(series_name, tsw) {
  if (series_name == "level_shift") {
    return(min(70, nrow(tsw)))
  }

  nrow(tsw)
}

collect_window_profiles <- function(y, series_name) {
  tsw <- ts_data(y, 12)
  anchor <- anchor_row(series_name, tsw)
  out <- data.frame()

  for (op_name in names(ops)) {
    preproc <- fit(ops[[op_name]], tsw)
    yt <- transform(preproc, tsw)
    yw <- yt[anchor, , drop = FALSE]

    out <- rbind(
      out,
      data.frame(
        lag = factor(colnames(yt), levels = colnames(yt)),
        value = as.vector(yw),
        operator = op_name,
        series = series_name
      )
    )
  }

  out
}

collect_t0_transforms <- function(y, series_name) {
  tsw <- ts_data(y, 12)
  out <- data.frame()

  for (op_name in names(ops)) {
    preproc <- ops[[op_name]]
    preproc <- fit(preproc, tsw)
    yt <- transform(preproc, tsw)
    out <- rbind(
      out,
      data.frame(
        idx = seq_len(nrow(yt)),
        value = as.vector(yt[, ncol(yt)]),
        operator = op_name,
        series = series_name
      )
    )
  }

  out
}

window_profiles <- do.call(
  rbind,
  Map(collect_window_profiles, synthetic, names(synthetic))
)

comparison_t0 <- do.call(
  rbind,
  Map(collect_t0_transforms, synthetic, names(synthetic))
)
```

This plot compares one fixed window per series after the full adaptive
normalization pipeline has been applied, including the internal global min-max
scaling. It is the clearest way to see how each operator transforms the same
lag profile while keeping the vertical scale comparable.

```{r}
ggplot(window_profiles, aes(x = lag, y = value, color = operator, group = operator)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.2) +
  facet_wrap(~ series, ncol = 2) +
  theme_minimal(base_size = 14)
```

### What To Look For In The Fixed-Window Plot

**Near-zero regime**

- `divide` can become very aggressive because small adaptive levels amplify the ratio
- `subtract` stays stable because it only removes the local level
- `softdivide` should stay closer to `subtract` here than to `divide`

**Level-shift regime**

- the main question is how the operators encode an abrupt change in local level
- `divide` and `softdivide` can look similar if the level term dominates the reference scale

**Heteroscedastic regime**

- the key question is whether similar oscillations at different amplitudes become more comparable
- `subtract` removes level but does not directly normalize scale
- `divide` and `softdivide` should compress amplitude differences more strongly

**Mixed regime**

- this combines drift, oscillation, and changing variance
- it is the best block to inspect overall trade-offs rather than a single ideal behavior

This second plot compares the transformed target dimension (`t0`) across time.
It is still useful, but it becomes easier to read after seeing the fixed-window
comparison above.

```{r}
ggplot(comparison_t0, aes(x = idx, y = value, color = operator)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ series, ncol = 2) +
  theme_minimal(base_size = 14)
```

### What To Look For In The `t0` Plot

- This plot shows how the final supervised target changes over time after normalization.
- It is useful for understanding temporal stability, but it is less direct than the fixed-window comparison.
- If two operators look close here, that does not necessarily mean they are identical. It may only mean that, for that series and that time range, both induce similar transformed targets.

### Why Some Curves Can Look Similar

- `divide` and `softdivide` can look close when the adaptive level dominates the local scale term. In that regime, `softdivide` approaches divisive behavior.
- `softdivide` and `asinh` can also look close when the normalized deviations stay in a moderate range, because `asinh(z)` behaves almost linearly around the origin.
- The `near_zero` and `heteroscedastic` blocks are usually the most informative ones for separating the four operators visually.

## Reading Summary

If you want a quick reading of the comparison, focus on these questions:

1. Which operator stays stable when the local level is close to zero?
2. Which operator best preserves comparability across changing amplitudes?
3. Which operator produces the smoothest compromise between additive and relative behavior?

In most practical cases:

- `subtract` is the safest additive baseline
- `divide` is the strongest relative normalizer, but it is fragile near zero
- `softdivide` is the main compromise operator
- `asinh` is the smooth nonlinear alternative when you want a bridge between additive and multiplicative interpretations

References

- Ogasawara, E., Martinez, L. C., De Oliveira, D., Zimbrão, G., Pappa, G. L., Mattoso, M. (2010).
Adaptive Normalization: A novel data normalization approach for non-stationary time series.
Proceedings of the International Joint Conference on Neural Networks (IJCNN).
doi:10.1109/IJCNN.2010.5596746

- Huber, P. J. (1964). Robust Estimation of a Location Parameter.
Annals of Mathematical Statistics, 35(1), 73-101. doi:10.1214/aoms/1177703732

- Burbidge, J. B., Magee, L., Robb, A. L. (1988). Alternative Transformations to Handle Extreme Values of the Dependent Variable.
Journal of the American Statistical Association, 83(401), 123-127.
