Sliding window min-max normalization: For each sliding window (row), compute its local min and max and rescale that window to [0,1]. This emphasizes local shape and pattern while removing level and amplitude differences across windows.

Objective: Apply per-window min-max normalization (each row scaled by its own minima and maxima) and visualize the effect of scaling.

```{r}
# Window-based Normalization (Min-Max)

# Installing the package (if needed)
#install.packages("tspredit")
```

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


```{r}
# Series for study

data(tsd)
```

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

```{r}
# Sliding windows

sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[,10])
```

```{r}
# Target (t0) visualization after windowing
library(ggplot2)
plot_ts(y=ts[,10]) + theme(text = element_text(size=16))
```

```{r}
# Normalization (fit and transform)

preproc <- ts_norm_swminmax()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[,10])
plot_ts(y=ts[1,]) + theme(text = element_text(size=16))
```

References
- C. M. Bishop (2006). Pattern Recognition and Machine Learning. Springer.
