Objective: Difference the series (first differences) and apply min-max normalization, observing level reduction and the effect on scale.

Method
First differencing transforms y_t to y_t − y_{t−1}, removing a unit‑root trend to stabilize the mean. After differencing, optional scaling (e.g., min‑max) standardizes the range for learning algorithms. Always compute scaling parameters on the training segment only.

```{r}
# Normalization by Differences (Diff)

# Installing the package (if needed)
#install.packages("tspredit")
```

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


```{r}
# Series for study

data(tsd)
```

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

```{r}
# Sliding windows

sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[,10])
```

```{r}
# Target (t0) visualization after windowing
library(ggplot2)
plot_ts(y=ts[,10]) + theme(text = element_text(size=16))
```

```{r}
# Normalization (fit and transform)

preproc <- ts_norm_diff()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[,9])
plot_ts(y=ts[,9]) + theme(text = element_text(size=16))
```

References
- G. E. P. Box, G. M. Jenkins, G. C. Reinsel, and G. M. Ljung (2015). Time Series Analysis: Forecasting and Control. Wiley.
