Adaptive normalization: AN centers and scales each window or time point using time‑varying statistics, typically computed via moving averages or exponentially weighted averages. This mitigates slow level/scale drift so models focus on local shape rather than absolute magnitude.

Objective: Apply adaptive normalization (moving average/EMA) to handle distribution shifts over time and observe the effect on the target.


```{r}
# Exponential Adaptive Normalization

# Installing the package (if needed)
#install.packages("tspredit")
```

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


```{r}
# Series for study

data(tsd)
```

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

```{r}
# Sliding windows

sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[,10])
```

```{r}
# Target (t0) visualization after windowing
library(ggplot2)
plot_ts(y=ts[,10]) + theme(text = element_text(size=16))
```

```{r}
# Normalization

preproc <- ts_norm_an()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[,10])
plot_ts(y=ts[1,]) + theme(text = element_text(size=16))
```

References
- Ogasawara, E., Martinez, L. C., De Oliveira, D., Zimbrão, G., Pappa, G. L., Mattoso, M. (2010).
Adaptive Normalization: A novel data normalization approach for non-stationary time series.
Proceedings of the International Joint Conference on Neural Networks (IJCNN).
doi:10.1109/IJCNN.2010.5596746
