Global Min-max: Compute the global minimum and maximum on the training data and rescale each value as (x − min) / (max − min). Apply the same parameters to validation/test to avoid leakage. This preserves relative ordering and maps all features to a common range.

Objective: Apply global min-max normalization (0 to 1) over the training series, transform the data, and visualize the effect of scaling.


```{r}
# Global Min-Max Normalization

# Installing the package (if needed)
#install.packages("tspredit")
```

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


```{r}
# Series for study

data(tsd)
```

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

```{r}
# Sliding windows

sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
summary(ts[,10])
```

```{r}
# Target (t0) visualization after windowing
library(ggplot2)
plot_ts(y=ts[,10]) + theme(text = element_text(size=16))
```

```{r}
# Normalization (fit and transform)

preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
ts_head(tst, 3)
summary(tst[,10])
plot_ts(y=ts[,10]) + theme(text = element_text(size=16))
```

References
- C. M. Bishop (2006). Pattern Recognition and Machine Learning. Springer.
