## ARIMA Baseline

About the method
- ARIMA models a univariate series with autoregressive terms, differencing, and moving-average terms.
- In `tspredit`, `ts_arima()` performs automatic order selection and then generates recursive forecasts for the requested horizon.

Didactic goal: learn the classical forecasting baseline before moving to sliding-window machine-learning models.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
#install.packages("tspredit")

# Loading the package
library(tspredit) 
```

We load the example series in its raw sequential form, without creating sliding windows yet.

```{r}
# Series for study (no sliding window)

data(tsd)
ts <- ts_data(tsd$y, 0)
ts_head(ts, 3)
```

Before moving on, we visualize the series so the effect of the next transformation can be compared against the original signal.

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

We now preserve the time order and split the raw sequence into train and test partitions.

```{r}
# Train-test split on the ordered series

samp <- ts_sample(ts, test_size = 5)
ts_head(samp$train, 3)
ts_head(samp$test, 3)
```

This step training the arima model (orders selected automatically).

```{r}
# Training the ARIMA model (orders selected automatically)

model <- ts_arima(p = 5, d = 0, q = 0)
set_example_seed()
model <- fit(model, x = samp$train)
```

We first evaluate the in-sample fit so the model adjustment can be compared with the later forecast.

```{r}
# Fit evaluation (train)

adjust <- predict(model, samp$train)
adjust <- as.vector(adjust)
output <- as.vector(samp$train)
ev_adjust <- evaluate(model, output, adjust)
ev_adjust$mse
```

We now generate a five-step forecast on the test segment and compare it with the observed values.

```{r}
# Forecast on test set (5 steps ahead)

prediction <- predict(model, x = samp$test[1,], steps_ahead = 5)
prediction <- as.vector(prediction)
output <- as.vector(samp$test)
ev_test <- evaluate(model, output, prediction)
ev_test
```

The final plot compares the observed series, the training adjustment, and the forecasted test horizon.

```{r}
# Plot comparing actual vs fit (train) and forecast (test)

yvalues <- c(samp$train, samp$test)
plot_ts_pred(y=yvalues, yadj=adjust, ypre=prediction, color_prediction="orange") + theme(text = element_text(size=16))
```

References
- G. E. P. Box, G. M. Jenkins, G. C. Reinsel, and G. M. Ljung (2015). Time Series Analysis: Forecasting and Control. Wiley.
- R. J. Hyndman and Y. Khandakar (2008). Automatic time series forecasting: The forecast package for R. Journal of Statistical Software, 27(3), 1–22. doi:10.18637/jss.v027.i03

