## Random Forest Regression

About the method
- Random forests combine many decision trees trained on bootstrap samples and random feature subsets.
- In forecasting with lagged windows, they capture nonlinear interactions among lags without requiring explicit parametric assumptions.

Didactic goal: inspect an ensemble baseline that is often strong with little manual feature engineering.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Time Series Regression - Random Forest

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


We load the example series that will be used throughout the demonstration.

```{r}
# Series for study and sliding windows

data(tsd)
ts <- ts_data(tsd$y, 10)
ts_head(ts, 3)
```

Before moving on, we visualize the series so the effect of the next transformation can be compared against the original signal.

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

We now preserve the time order, split the data into train and test partitions, and project the windows into inputs and targets.

```{r}
# Train-test split and projection (X, y)

samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test <- ts_projection(samp$test)
```

We now prepare the preprocessing object that will normalize the inputs before the model is trained.

```{r}
# Preprocessing (global min-max normalization)

preproc <- ts_norm_gminmax()
```

We now train the random forest model on the prepared training data. For recursive forecasting, a slightly larger lag set and a less tiny ensemble usually behave better than the smallest possible defaults.

```{r}
# Training the Random Forest model

model <- ts_rf(ts_norm_gminmax(), input_size = 9, nodesize = 1, ntree = 100)
set_example_seed()
model <- fit(model, x=io_train$input, y=io_train$output)
```

If you later tune this model, a good first search space is usually:

- `input_size` near the most recent 6 to 10 lags when the sliding window has size 10;
- `nodesize` in small values such as `c(1, 3, 5)`;
- `ntree` in moderate ensembles such as `c(50, 100, 200)`;
- `mtry` in a small range like `2:5`, always respecting the chosen `input_size`.

We first evaluate the in-sample fit so the model adjustment can be compared with the later forecast.

```{r}
# Fit evaluation (train)

adjust <- predict(model, io_train$input)
adjust <- as.vector(adjust)
output <- as.vector(io_train$output)
ev_adjust <- evaluate(model, output, adjust)
ev_adjust$mse
```

We now generate a five-step forecast on the test segment and compare it with the observed values.

```{r}
# Forecast on test set (5 steps ahead)

prediction <- predict(model, x=io_test$input[1,], steps_ahead=5)
prediction <- as.vector(prediction)
output <- as.vector(io_test$output)
ev_test <- evaluate(model, output, prediction)
ev_test
```

The final plot compares the observed series, the training adjustment, and the forecasted test horizon.

```{r}
# Plot comparing actual vs fit (train) and forecast (test)

yvalues <- c(io_train$output, io_test$output)
plot_ts_pred(y=yvalues, yadj=adjust, ypre=prediction, color_prediction="orange") + theme(text = element_text(size=16))
```

References
- L. Breiman (2001). Random Forests. Machine Learning, 45(1), 5–32.

