## ELM Regression

About the method
- Extreme Learning Machines use a single hidden layer with randomly assigned hidden weights and a fast output-layer fit.
- They are attractive when you want a neural model with much cheaper training than iterative backpropagation.

Didactic goal: compare a fast neural regressor with the more conventional MLP workflow.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Time Series Regression - ELM

# Installing the package (if needed)
#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
# Loading the packages
library(daltoolbox)
library(tspredit) 
```


We load the example series that will be used throughout the demonstration.

```{r}
# Series for study and sliding windows

data(tsd)
ts <- ts_data(tsd$y, 10)
ts_head(ts, 3)
```

Before moving on, we visualize the series so the effect of the next transformation can be compared against the original signal.

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

We now preserve the time order, split the data into train and test partitions, and project the windows into inputs and targets.

```{r}
# Train-test split and projection (X, y)

samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test <- ts_projection(samp$test)
```

We now prepare the preprocessing object that will normalize the inputs before the model is trained.

```{r}
# Preprocessing (global min-max normalization)

preproc <- ts_norm_gminmax()
```

We now train the elm model on the prepared training data.

```{r}
# Training the ELM model

model <- ts_elm(ts_norm_gminmax(), input_size=4, nhid=3, actfun="purelin")
set_example_seed()
model <- fit(model, x=io_train$input, y=io_train$output)
```

We first evaluate the in-sample fit so the model adjustment can be compared with the later forecast.

```{r}
# Fit evaluation (train)

adjust <- predict(model, io_train$input)
adjust <- as.vector(adjust)
output <- as.vector(io_train$output)
ev_adjust <- evaluate(model, output, adjust)
ev_adjust$mse
```

We now forecast the test set and compare the predicted values with the observed ones.

```{r}
# Forecast on test set

prediction <- predict(model, x=io_test$input[1,], steps_ahead=5)
prediction <- as.vector(prediction)
output <- as.vector(io_test$output)
ev_test <- evaluate(model, output, prediction)
ev_test
```

This final plot summarizes the result of the transformation so the effect can be interpreted visually.

```{r}
# Plot results

yvalues <- c(io_train$output, io_test$output)
plot_ts_pred(y=yvalues, yadj=adjust, ypre=prediction, color_prediction="orange") + theme(text = element_text(size=16))
```

References
- G.-B. Huang, Q.-Y. Zhu, and C.-K. Siew (2006). Extreme learning machine: Theory and applications. Neurocomputing, 70(1–3), 489–501.

