## Conv1D Regression

About the method
- A 1D convolutional network applies learnable filters across the lag window to detect local temporal motifs.
- It is useful when short repeated patterns matter more than long global recurrence alone.

Didactic goal: understand how convolution changes the representation learned from sliding windows.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Time Series Regression - 1D CNN (Conv1D)

# Installing packages (if needed)

#install.packages("tspredit")
```

We start by loading the packages used throughout this example.

```{r}
# Loading the packages
library(daltoolbox)
library(daltoolboxdp)
library(tspredit)
```

We load the example series that will be used throughout the demonstration.

```{r}
# Series for study and sliding windows

data(tsd)
ts <- ts_data(tsd$y, 10)
ts_head(ts, 3)
```

Before moving on, we visualize the series so the effect of the next transformation can be compared against the original signal.

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

We now preserve the time order, split the data into train and test partitions, and project the windows into inputs and targets.

```{r}
# Train-test split and projection (X, y)

samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test <- ts_projection(samp$test)
```

This step training the 1d cnn.

```{r}
# Training the 1D CNN

model <- ts_conv1d(ts_norm_gminmax(), input_size=4, epochs=10000)
set_example_seed()
model <- fit(model, x=io_train$input, y=io_train$output)
```

We first evaluate the in-sample fit so the model adjustment can be compared with the later forecast.

```{r}
# Fit evaluation (train)

adjust <- predict(model, io_train$input)
adjust <- as.vector(adjust)
output <- as.vector(io_train$output)
ev_adjust <- evaluate(model, output, adjust)
ev_adjust$mse
```

We now forecast the test set and compare the predicted values with the observed ones.

```{r}
# Forecast on test set

prediction <- predict(model, x=io_test$input[1,], steps_ahead=5)
prediction <- as.vector(prediction)
output <- as.vector(io_test$output)
ev_test <- evaluate(model, output, prediction)
ev_test
```

This final plot summarizes the result of the transformation so the effect can be interpreted visually.

```{r}
# Plot results

yvalues <- c(io_train$output, io_test$output)
plot_ts_pred(y=yvalues, yadj=adjust, ypre=prediction, color_prediction="orange") + theme(text = element_text(size=16))
```

References
- Y. LeCun, L. Bottou, Y. Bengio, and P. Haffner (1998). Gradient-based learning applied to document recognition. Proceedings of the IEEE, 86(11), 2278–2324.

