Conv1D: One-dimensional Convolutional Neural Networks (1D CNNs) extract local temporal patterns by convolving learnable filters across the input window. Stacked convolution and pooling layers can capture increasingly abstract features before a regression head maps to the next-step forecast. CNNs are effective when short- to mid-range motifs repeat over time and benefit from normalized inputs.

Objective: Train and evaluate a 1D CNN (Conv1D) for time-series forecasting with sliding windows, including normalization, fitting, and evaluation.

```{r}
# Time Series Regression - 1D CNN (Conv1D)

# Installing packages (if needed)

#install.packages("tspredit")
```

```{r}
# Loading the packages
library(daltoolbox)
library(daltoolboxdp)
library(tspredit)
```

```{r}
# Series for study and sliding windows

data(tsd)
ts <- ts_data(tsd$y, 10)
ts_head(ts, 3)
```

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

```{r}
# Train-test split and projection (X, y)

samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test <- ts_projection(samp$test)
```

```{r}
# Training the 1D CNN

model <- ts_conv1d(ts_norm_gminmax(), input_size=4, epochs=10000)
model <- fit(model, x=io_train$input, y=io_train$output)
```

```{r}
# Fit evaluation (train)

adjust <- predict(model, io_train$input)
adjust <- as.vector(adjust)
output <- as.vector(io_train$output)
ev_adjust <- evaluate(model, output, adjust)
ev_adjust$mse
```

```{r}
# Forecast on test set

prediction <- predict(model, x=io_test$input[1,], steps_ahead=5)
prediction <- as.vector(prediction)
output <- as.vector(io_test$output)
ev_test <- evaluate(model, output, prediction)
ev_test
```

```{r}
# Plot results

yvalues <- c(io_train$output, io_test$output)
plot_ts_pred(y=yvalues, yadj=adjust, ypre=prediction) + theme(text = element_text(size=16))
```

References
- Y. LeCun, L. Bottou, Y. Bengio, and P. Haffner (1998). Gradient-based learning applied to document recognition. Proceedings of the IEEE, 86(11), 2278–2324.
