## Build `ts_data`

About the technique
- `ts_data()` converts a raw time series into the tabular representation used by `tspredit`.
- Each row stores one forecasting context and the lag columns (`t{k}`) define what the model can observe before predicting the next value.

Didactic goal: understand the internal representation that later projection, preprocessing, and prediction examples rely on.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
#install.packages("tspredit")

# Loading the package
library(tspredit) 
```

We load the example data and separate the variables that will be used in the following steps.

```{r}
# Load data and split axis/value
data(tsd)
x <- tsd$x
y <- tsd$y
```

We plot the data here so the effect of the next transformation can be compared visually.

```{r}
library(ggplot2)
# Visualize the original series
plot_ts(x = x, y = y) + theme(text = element_text(size=16))
```

We first build `ts_data` without a sliding window so the base structure is easier to inspect.

```{r}
# ts_data without window (only t0)
data <- ts_data(y)
ts_head(data)
plot_ts(y=data[,1]) + theme(text = element_text(size=16))
```

We now rebuild the same series with a sliding window of size 10 so each row stores recent lags explicitly.

```{r}
# ts_data with window size 10 (t9..t0)
data10 <- ts_data(y, 10)
ts_head(data10)
```

This indexing example select a row, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a row
r1 <- data10[12,]
r1
```

This indexing example select a range of rows, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a range of rows
r2 <- data10[12:13,]
r2
```

This indexing example select a column, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a column
c1 <- data10[,1]
ts_head(c1)
```

This indexing example select a range of columns, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a range of columns
c2 <- data10[,1:2]
ts_head(c2)
```

This indexing example select a range of rows and columns, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a range of rows and columns
rc1 <- data10[12:13,1:2]
rc1
```

This indexing example select one row and a range of columns, so you can inspect the structure returned by `ts_data`.

```{r}
# Select one row and a range of columns
rc2 <- data10[12,1:2]
rc2
```

This indexing example select a range of rows and one column, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a range of rows and one column
rc3 <- data10[12:13,1]
rc3
```

This indexing example select a single observation, so you can inspect the structure returned by `ts_data`.

```{r}
# Select a single observation
rc4 <- data10[12,1]
rc4
```
