## Project `ts_data` into `X` and `y`

About the technique
- `ts_projection()` separates the lagged attributes from the forecast target.
- This is the bridge between the time-series representation and the supervised-learning interface expected by forecasting models.

Didactic goal: see exactly how a time-series window becomes model input (`X`) and expected output (`y`).

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
#install.packages("tspredit")

# Loading the package
library(tspredit) 
```

We load the example series that will be used throughout the demonstration.

```{r}
# Series for study

data(tsd)
```

Before moving on, we visualize the series so the effect of the next transformation can be compared against the original signal.

```{r}
# Series visualization
library(ggplot2)
plot_ts(x=tsd$x, y=tsd$y) + theme(text = element_text(size=16))
```

The next step organizes the series into sliding windows, which is the tabular representation used by the later transformations and models.

```{r}
# Sliding windows

sw_size <- 5
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
```

We now preserve the time order, split the data into train and test partitions, and project the windows into inputs and targets.

```{r}
# Projection (X, y)

io <- ts_projection(ts)
```

This chunk input data (x).

```{r}
# Input data (X)
ts_head(io$input)
```

This chunk output data (y).

```{r}
# Output data (y)
ts_head(io$output)
```
