## Time-Ordered Train/Test Split

About the technique
- `ts_sample()` creates training and test segments without shuffling the series.
- In forecasting, preserving chronology is essential because the model must learn from the past and be evaluated on future observations only.

Didactic goal: understand how `tspredit` defines a valid forecasting split before any model is fitted.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
#install.packages("tspredit")

# Loading the package
library(tspredit) 
```

We load the example series that will be used throughout the demonstration.

```{r}
# Series for study

data(tsd)
```

We plot the data here so the effect of the next transformation can be compared visually.

```{r}
library(ggplot2)
plot_ts(x = tsd$x, y = tsd$y) + theme(text = element_text(size=16))
```

The next step organizes the series into sliding windows, which is the tabular representation used by the later transformations and models.

```{r}
# Sliding windows

sw_size <- 10
ts <- ts_data(tsd$y, sw_size)
ts_head(ts, 3)
```

This chunk sampling (train and test).

```{r}
# Sampling (train and test)

test_size <- 3
samp <- ts_sample(ts, test_size)
```

This chunk first five rows of the train set.

```{r}
# First five rows of the train set
ts_head(samp$train, 5)
```

This chunk last five rows of the train set.

```{r}
# Last five rows of the train set
ts_head(samp$train[-c(1:(nrow(samp$train)-5)),])
```

This chunk test data.

```{r}
# Test data
ts_head(samp$test)
```
