---
title: "ts_preprocessing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ts_preprocessing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# DAL Package.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(daltoolbox)
```

## Time series preprocessing

### Cosine series for studying
```{r}
i <- seq(0, 2*pi, pi/50)
x <- cos(i)

plot(i, x)
lines(i, x)
```

### sliding windows
```{r}
sw_size <- 10
ts <- ts_data(x, sw_size)
tshead(ts, 3)
summary(ts[,10])
plot(ts[,10])
```

### global min-max
```{r}
preproc <- ts_gminmax()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,10])
plot(tst[,10])
```

### global min-max (differentiation)
```{r}
preproc <- ts_diff()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,9])
plot(tst[,9])
```

### sliding window min-max
```{r}
preproc <- ts_swminmax()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,10])
#ploting the first window
plot(as.vector(tst[1,]))
```

### sliding window adaptive normalization
```{r}
preproc <- ts_an()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,10])
#ploting the first window
plot(as.vector(tst[1,]))
```

### sliding window adaptive normalization
Using nw parameter
```{r}
preproc <- ts_an(nw=3)
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,10])
#ploting the first window
plot(as.vector(tst[1,]))
```

### sliding window adaptive normalization (exponential moving average)
```{r}
preproc <- ts_ean()
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,10])
#ploting the first window
plot(as.vector(tst[1,]))
```

### sliding window adaptive normalization (exponential moving average)
using nw parameter
```{r}
preproc <- ts_ean(nw = 3)
preproc <- fit(preproc, ts)
tst <- transform(preproc, ts)
tshead(tst, 3)
summary(tst[,10])
#ploting the first window
plot(as.vector(tst[1,]))
```
