---
title: "ts_tune"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ts_tune}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# DAL Package

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(daltoolbox)
library(elmNNRcpp)
```

## Time Series regression - tune

### Cosine series for studying
```{r}
i <- seq(0, 25, 0.25)
x <- cos(i)

plot(i, x)
lines(i, x)
```

### sliding windows
```{r}
sw_size <- 10
ts <- ts_data(x, sw_size)
tshead(ts, 3)
```

### data sampling
```{r}
test_size <- 1
samp <- ts_sample(ts, test_size)
tshead(samp$train, 3)
tshead(samp$test)
```

### Model training

Example is using ts_elm() as base_model, but all time series models can be used. 

Options for ranges for all time series models are presented in the end of this notebook.

Options for the list of preprocessing are presented in the end of this notebook.

Options for the list of data augmentation are presented in the end of this notebook.

Input size options should be between 1 and sw_size-2

```{r}
tune <- ts_maintune(preprocess=list(ts_gminmax()), input_size=c(3:5), base_model = ts_elm(), augment = list(ts_augment()))
ranges <- list(nhid = 1:5, actfun=c('sig', 'radbas', 'tribas', 'relu', 'purelin'))

io_train <- ts_projection(samp$train)
model <- fit(tune, x=io_train$input, y=io_train$output, ranges)

print(describe(model))
```

### Evaluation of adjustment
```{r}
adjust <- predict(model, io_train$input)
ev_adjust <- evaluation.tsreg(io_train$output, adjust)
print(head(ev_adjust$metrics))
```

### Prediction of test
```{r}
steps_ahead <- 1
io_test <- ts_projection(samp$test)
prediction <- predict(model, x=io_test$input, steps_ahead=steps_ahead)
prediction <- as.vector(prediction)

output <- as.vector(io_test$output)
if (steps_ahead > 1)
    output <- output[1:steps_ahead]

print(sprintf("%.2f, %.2f", output, prediction))
```

### Evaluation of test data
```{r}
ev_test <- evaluation.tsreg(output, prediction)
print(head(ev_test$metrics))
print(sprintf("%s: smape: %.2f", describe(model), 100*ev_test$metrics$smape))
```

---

### Otions for machine learning
```R
### Ranges for ELM
ranges_elm <- list(nhid = 1:20, actfun=c('sig', 'radbas', 'tribas', 'relu', 'purelin'))

### Ranges for MLP
ranges_mlp <- list(size = 1:10, decay = seq(0, 1, 1/9), maxit=10000)

### Ranges for RF
ranges_rf <- list(nodesize=1:10, ntree=1:10)

### Ranges for SVM
ranges_svm <- list(kernel=c("radial", "poly", "linear", "sigmoid"), epsilon=seq(0, 1, 0.1), cost=seq(20, 100, 20))

### Ranges for LSTM
ranges_lstm <- list(epochs=10000)
```

### Options for preprocessing
```R
preprocessing <- list(ts_gminmax(), ts_diff(), ts_swminmax(), ts_an(), ts_ean())
```

### Options for data augmentation
```R
augment <- list(ts_augment(), jitter(), stretch(), shrink(), flip(), wormhole())
```




