## Lag Mapping Comparison

About the comparison
- Different lag selectors emphasize different assumptions about temporal dependence.
- This notebook compares the resulting lag subsets on synthetic signals that isolate short-memory, seasonal, and multiscale behavior.

Didactic goal: visualize how the mapping families differ before they are coupled to a predictor.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/tspredit/main/examples/seed.R"))
# Lag mapping comparison
```

```{r}
library(daltoolbox)
library(tspredit)
library(ggplot2)
```

```{r}
set_example_seed()

n <- 120
x <- 1:n

series_short <- sin(x / 5) + rnorm(n, sd = 0.15)
series_seasonal <- sin(2 * pi * x / 12) + 0.4 * sin(2 * pi * x / 6) + rnorm(n, sd = 0.1)
series_multiscale <- 0.6 * sin(x / 4) + 0.3 * sin(x / 13) + 0.01 * x + rnorm(n, sd = 0.12)

synthetic <- list(
  short = series_short,
  seasonal = series_seasonal,
  multiscale = series_multiscale
)
```

```{r}
plot_ts(y = synthetic$short)
plot_ts(y = synthetic$seasonal)
plot_ts(y = synthetic$multiscale)
```

```{r}
methods <- c("recent", "even", "geom", "acf", "pacf", "seasonal", "mi", "mrmr")

describe_map <- function(series_name, series_values) {
  ts <- ts_data(series_values, 13)
  io <- ts_projection(ts)

  do.call(rbind, lapply(methods, function(method_name) {
    mapper <- ts_lagmap(method = method_name, seasonality = 12)
    mapper <- fit(mapper, io$input, io$output, input_size = 5)
    data.frame(
      series = series_name,
      method = method_name,
      lag = mapper$lags
    )
  }))
}

mapping_df <- do.call(
  rbind,
  Map(describe_map, names(synthetic), synthetic)
)

mapping_df
```

```{r}
ggplot(mapping_df, aes(x = lag, y = method, color = series)) +
  geom_point(size = 3) +
  facet_wrap(~series, ncol = 1) +
  scale_x_reverse() +
  labs(
    title = "Lag subsets chosen by different mapping rules",
    x = "Lag",
    y = "Mapping method"
  )
```
